Analytical choices for analyzing multidimensional behavior - Many analysts test hypotheses about human speech.

### 2021-01-27 RR_manuscript/Draft_RR.Rmd
:memo: - render RR + bib updates

Data analysis comes with many decisions like how to measure a given phenomenon or behavior, what data to submit to statistical modeling and which to exclude in the final analysis, or what inferential decision procedure to apply.
However, if these decisions during data analysis are not specified in advance, we might stumble upon seemingly meaningful patterns in the data that are merely statistical flukes.
<!-- I'd probably drop this. When we make unspecified decisions during data analysis we can stumble upon spurious results (most likely) as well as real results (less likely, still possible)... in other words it might not all be error... the issue is that we won't know the difference -->
<!-- TR: But isn't the human bias the major issue in statistical misinterpretation? (It also leads naturally to the next paragraph.) -->
This can be problematic because humans show cognitive biases that can lead to erroneous inferences.
Humans filter the world in irrational ways [e.g., @tversky1974judgment], seeing coherent patterns in randomness (Brugger 2001), convincing themselves of the validity of prior expectations ["I knew it", @nickerson1998confirmation], and perceiving events as being plausible in hindsight ["I knew it all along", @fischhoff1975hindsight]. <!-- JC: If helpful we could also give a concrete common example here, like regarding forecasting the weather or something -->
<!-- TR: Generally a good idea, but I'd say we keep it crisp at that point. The take-home here is: Humans are biased. -->
In connection with an academic incentive system that rewards certain discovery processes more than others [@sterling1959publication; @koole2012rewarding], we often find ourselves exploring many possible analytical pipelines, but only reporting a select few.
This issue is particularly amplified in fields in which the raw data lend themselves to many possible ways to measure [@roettger2019researcher].
Combined with a wide variety of methodological and theoretical traditions as well as varying levels of statistical training across subfields, the inherent flexibility of data analysis might lead to a vast plurality of analytic approaches that can lead to different scientific conclusions.
Consequently, there might be many published papers that present overconfident interpretations of their data based on idiosyncratic analytic strategies [e.g., @simmons2011false; @gelman2014statistical].
These interpretations are either associated with an unknown amount of uncertainty or lend themselves to alternative interpretation if analyzed differently.
However, instead of being critically evaluated, scientific results often remain unchallenged in the publication record.
Despite recent efforts to improve transparency and reproducibility (REFS) and freely available and accessible infrastructures such as provided by the Open Science Framework (osf.io, ADD), critical re-analyses of published analytic strategies are still not very common because data sharing remains rare [@wicherts2006poor].

@@

Moving onto the expression of higher-level functions like information structure and discourse pragmatics, relevant acoustic cues can be distributed throughout even larger domains, such as phrases and whole utterances [e.g., @ladd2008intonational].
Differences in position, shape, and alignment of pitch modulations over multiple locations within a sentence are correlated with differences in discourse functions (e.g. Niebuhr et al., 2011).
The latter can also be expressed by global vs local pitch modulations (Haan 2002), as well as acoustic information within the temporal or spectral domain [e.g., @van2005speech].
Extra-linguistic information, like speaker's intentions, levels of arousal or social identity, are also conveyed by broad-domain parameters, such as voice quality, rhythm, and pitch [@foulkes2006; @ogden2004; @white2009].

@@

These decisions are usually made prior to any statistical analysis, but are at times revised a posteriori (i.e. after data collection and/or preliminary analyses) in light of unforeseen or surprising outcomes.
These myriads of possible decisions are exponentiated by researcher degrees of freedom related to statistical analysis (e.g. Wicherts et al. ????). <!-- year needed -->
Even the analysis of a single measure can be approached via an ever-increasing range of different statistical models (REFs).
The present paper probes this garden of forking paths in the analysis of speech.
To assess the variability in data analysis pipelines across independent researchers, we provided XX analytical teams with an experimentally elicited speech corpus and asked them to investigate acoustic differences related to a functional contrast that might be manifested across the whole utterance.

@@

To summarize the variability in reported effect sized, we will follow Bayesian random-effects meta-analytical techniques.
Based on the common practices currently in place within the field, we anticipate that researchers will use multi-level/hierarchical/random-effects regression models, so that common effect size measures such as Cohen's $d$ would be inappropriate.
<!-- NOTE: (SC) it would be good to decide on a clear terminology to refer to the individual original models, the individual models that we run (to refit the original models), and the meta-analytical model.  -->
Since the variables used by the analysis teams might substantially differ in their measurement scales (e.g, Hz for frequency vs ms for duration), we will standardize all reported effects by refitting each reported model with centered and scaled continuous variables (*z*-scores, i.e. the observed values subtracted from the mean divided by the standard deviation) and sum-coded factor variables.
\highlight[id=SC, comment = {Is this sentence good?}]{Factor-level ordering} for each factor variable will be decided on a model-by-model basis, depending on which levels were compared by the team.
Each standardized (refitted) model will be fitted as a Bayesian regression model with Stan [@stan-development-team2020], RStan [@stan-development-team2020a], and brms [@burkner2017] in R [@r-core-team2020].
For those reported models that were originally fitted within a frequentist approach, uniform distributions will be used as the priors of all parameters (with the restriction that only positive numbers will be included for scale parameters), making the standardized models in fact equivalent to the reported frequentist models.
If a team has fitted Bayesian models, the same priors as reported by the team will be used in fitting the respective standardized model.

@@

The outcome variable will be the set of standardized effect sizes $\eta_i$.
The likelihood of $\eta_i$ is assumed to correspond to a normal distribution [@knight2000].
The analysis teams will constitute the group-level effect (i.e., random effect, `(1 | team)`).
The standard errors $\text{se}_i$ will be included as the standard deviation $\sigma_i$ of $\eta_i$ to fit a measurement-error model, as discussed above.
We will use regularizing weakly-informative priors for the intercept $\alpha$ ($Normal(0, 1)$) and for the group-level standard deviation $\alpha_{\text{t}[i]}$ ($HalfCauchy(0, 1)$).
<!-- TR: isn't an SD of 1 rather conservative. Also small sample size, maybe rather using cauchy for the thick tails? -->
<!-- SC: For the pop-level params I would stick to the Normal, or if we want to avoid problems with thick tails I'd go with a student-t. -->
We will fit this model with 4 chains of Hamiltonian Monte-Carlo sampling for the estimation of the joint posterior distribution, using the No U-Turn Sampler (NUTS) as implemented in Stan, and 4000 iterations (2000 for warm-up) per chain, distributed across 4 processing cores.
\highlight[id=SC, comment = {This ok?}]{In case of issues from divergent transitions} in the sampling, we will increase `adapt_delta`, `tree_depth`, and the number of iterations in this order until there are no divergent transitions.
The analysis will be conducted in R [@r-core-team2020] and fit using Stan [@stan-development-team2020], RStan [@stan-development-team2020a], and brms [@burkner2017].
<!-- TODO: add link to code of the meta-analytical model -->
The code used to run the model can be found here: INSERT LINK.

>>>>

Data analysis comes with many decisions like how to measure a given phenomenon or behavior, what data to submit to statistical modeling and which to exclude in the final analysis, or what inferential decision procedure to apply.
However, if these decisions during data analysis are not specified in advance, we might stumble upon seemingly meaningful patterns in the data that are merely statistical flukes.
<!-- I'd probably drop this. When we make unspecified decisions during data analysis we can stumble upon spurious results (most likely) as well as real results (less likely, still possible)... in other words it might not all be error... the issue is that we won't know the difference -->
<!-- TR: But isn't the human bias the major issue in statistical misinterpretation? (It also leads naturally to the next paragraph.) -->
This can be problematic because humans show cognitive biases that can lead to erroneous inferences.
Humans filter the world in irrational ways [e.g., @tversky1974judgment], seeing coherent patterns in randomness [@brugger2001], convincing themselves of the validity of prior expectations ["I knew it", @nickerson1998confirmation], and perceiving events as being plausible in hindsight ["I knew it all along", @fischhoff1975hindsight]. <!-- JC: If helpful we could also give a concrete common example here, like regarding forecasting the weather or something -->
<!-- TR: Generally a good idea, but I'd say we keep it crisp at that point. The take-home here is: Humans are biased. -->
In connection with an academic incentive system that rewards certain discovery processes more than others [@sterling1959publication; @koole2012rewarding], we often find ourselves exploring many possible analytical pipelines, but only reporting a select few.
This issue is particularly amplified in fields in which the raw data lend themselves to many possible ways to measure [@roettger2019researcher].
Combined with a wide variety of methodological and theoretical traditions as well as varying levels of statistical training across subfields, the inherent flexibility of data analysis might lead to a vast plurality of analytic approaches that can lead to different scientific conclusions.
Consequently, there might be many published papers that present overconfident interpretations of their data based on idiosyncratic analytic strategies [e.g., @simmons2011false; @gelman2014statistical].
These interpretations are either associated with an unknown amount of uncertainty or lend themselves to alternative interpretation if analyzed differently.
However, instead of being critically evaluated, scientific results often remain unchallenged in the publication record.
Despite recent efforts to improve transparency and reproducibility (REFS) and freely available and accessible infrastructures such as provided by the Open Science Framework (osf.io, ADD), critical re-analyses of published analytic strategies are still not very common because data sharing remains rare [@wicherts2006poor].

@@

Moving onto the expression of higher-level functions like information structure and discourse pragmatics, relevant acoustic cues can be distributed throughout even larger domains, such as phrases and whole utterances [e.g., @ladd2008intonational].
Differences in position, shape, and alignment of pitch modulations over multiple locations within a sentence are correlated with differences in discourse functions (e.g. Niebuhr et al., 2011).
The latter can also be expressed by global vs local pitch modulations [@heuven2002], as well as acoustic information within the temporal or spectral domain [e.g., @van2005speech].
Extra-linguistic information, like speaker's intentions, levels of arousal or social identity, are also conveyed by broad-domain parameters, such as voice quality, rhythm, and pitch [@foulkes2006; @ogden2004; @white2009].

@@

These decisions are usually made prior to any statistical analysis, but are at times revised a posteriori (i.e. after data collection and/or preliminary analyses) in light of unforeseen or surprising outcomes.
These myriads of possible decisions are exponentiated by researcher degrees of freedom related to statistical analysis [e.g. wicherts2016].
Even the analysis of a single measure can be approached via an ever-increasing range of different statistical models (REFs).
The present paper probes this garden of forking paths in the analysis of speech.
To assess the variability in data analysis pipelines across independent researchers, we provided XX analytical teams with an experimentally elicited speech corpus and asked them to investigate acoustic differences related to a functional contrast that might be manifested across the whole utterance.

@@

To summarize the variability in reported effect sized, we will follow Bayesian random-effects meta-analytical techniques.
Based on the common practices currently in place within the field, we anticipate that researchers will use multi-level/hierarchical/random-effects regression models, so that common effect size measures such as Cohen's $d$ would be inappropriate.
<!-- NOTE: (SC) it would be good to decide on a clear terminology to refer to the individual original models, the individual models that we run (to refit the original models), and the meta-analytical model.  -->
Since the variables used by the analysis teams might substantially differ in their measurement scales (e.g, Hz for frequency vs ms for duration), we will standardize all reported effects by refitting each reported model with centered and scaled continuous variables (*z*-scores, i.e. the observed values subtracted from the mean divided by the standard deviation) and sum-coded factor variables.
\highlight[id=SC, comment = {Is this sentence good?}]{Factor-level ordering} for each factor variable will be decided on a model-by-model basis, depending on which levels were compared by the team.
Each standardized (refitted) model will be fitted as a Bayesian regression model with Stan [@stan2021], RStan [@stan2020a], and brms [@burkner2017] in R [@R-base].
For those reported models that were originally fitted within a frequentist approach, uniform distributions will be used as the priors of all parameters (with the restriction that only positive numbers will be included for scale parameters), making the standardized models in fact equivalent to the reported frequentist models.
If a team has fitted Bayesian models, the same priors as reported by the team will be used in fitting the respective standardized model.

@@

The outcome variable will be the set of standardized effect sizes $\eta_i$.
The likelihood of $\eta_i$ is assumed to correspond to a normal distribution [@knight2000].
The analysis teams will constitute the group-level effect (i.e., random effect, `(1 | team)`).
The standard errors $\text{se}_i$ will be included as the standard deviation $\sigma_i$ of $\eta_i$ to fit a measurement-error model, as discussed above.
We will use regularizing weakly-informative priors for the intercept $\alpha$ ($Normal(0, 1)$) and for the group-level standard deviation $\alpha_{\text{t}[i]}$ ($HalfCauchy(0, 1)$).
<!-- TR: isn't an SD of 1 rather conservative. Also small sample size, maybe rather using cauchy for the thick tails? -->
<!-- SC: For the pop-level params I would stick to the Normal, or if we want to avoid problems with thick tails I'd go with a student-t. -->
We will fit this model with 4 chains of Hamiltonian Monte-Carlo sampling for the estimation of the joint posterior distribution, using the No U-Turn Sampler (NUTS) as implemented in Stan, and 4000 iterations (2000 for warm-up) per chain, distributed across 4 processing cores.
\highlight[id=SC, comment = {This ok?}]{In case of issues from divergent transitions} in the sampling, we will increase `adapt_delta`, `tree_depth`, and the number of iterations in this order until there are no divergent transitions.
The analysis will be conducted in R [@R-base] and fit using Stan [@stan2021], RStan [@stan2020a], and brms [@burkner2017].
<!-- TODO: add link to code of the meta-analytical model -->
The code used to run the model can be found here: INSERT LINK.
